---
title: "diagonals"
author: "Bastiaan Quast"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Block-diagonal matrices can be extracted or removed using two small functions implemented here, in addition, non-square matrices are supported.

Block diagonal matrices occur when we combine two dimensions of a data set along one edge of a matrix. For example, trade-flow data in the [decompr](http://cran.r-project.org/web/packages/decompr/index.html) and [gvc](http://cran.r-project.org/web/packages/gvc/index.html) package have each country-industry combination occur on each edge of the matrix.

```{r, matrix}
m <- matrix(1:64, nrow=8)
m
```

We can now drop everything except a block diagonal.

```{r, block_matrix}
library(diagonals)
block_matrix(m, steps=4)
```

Or drop only the block diagonal.

```{r, minus_block_matrix}
minus_block_matrix(m, steps=2)
```

The number of `steps` can vary, but dimensions should be multiples of `steps`. Alternatively we can specify the size of the `size` of the step.

```{r, steps}
block_matrix(m, size=2)
```

Even non-square matrices are supported, as long as each dimension is a multiple of size (i.e. `dim(matrix) %% size == c(0, 0)` ).

```{r non-squre}
nsm <- matrix(1:27, nrow=9, ncol=3 )
block_matrix(nsm, steps=3)
```

Or this.

```{r}
block_matrix( t(nsm), steps=3 )
```
