---
title: "diagonals"
author: "Bastiaan Quast"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{diagonals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Block-diagonal matrices can be extracted or removed using two small functions implemented here, in addition, non-square matrices are supported.

Block diagonal matrices occur when we combine two dimensions of a data set along one edge of a matrix. For example, trade-flow data in the [decompr](http://cran.r-project.org/web/packages/decompr/index.html) and [gvc](http://cran.r-project.org/web/packages/gvc/index.html) package have each country-industry combination occur on each edge of the matrix.

```{r, matrix}
m <- matrix(1:64, nrow=8)
m
```

We can now drop everything except a block diagonal.

```{r, block_matrix}
library(diagonals)
block_matrix(m, steps=4)
```

Or drop only the block diagonal.

```{r, minus_block_matrix}
minus_block_matrix(m, steps=2)
```

The number of `steps` can vary, but dimensions should be multiples of `steps`. Alternatively we can specify the size of the `size` of the step.

```{r, steps}
block_matrix(m, size=2)
```

Even non-square matrices are supported, as long as each dimension is a multiple of size.

```{r create-non-square}
# create the matrix
nsm <- matrix(1:27, nrow=9, ncol=3 )

# test if the dimensions are multiples of size
# i.e. there are no remainders after the modulo division of dimensions by steps
dim(nsm) %% 3 # we will use 3 (three) steps
```

Now extract the rectangular diagonal.

```{r apply-non-square}
rectangle_matrix(nsm, steps=3)
```

Or this.

```{r}
rectangle_matrix( t(nsm), steps=3 )
```

Lastly, we can also choose a replacement parameter other than zero.

```{r}
block_matrix(m, steps=4, replacement=-1)
```

or 

```{r}
minus_rectangle_matrix(nsm, steps=3, replacement=666)
```
