---
title: "Fat Diagonals"
author: "Bastiaan Quast"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fat Diagonals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Fat diagonal matrices occur when we combine two dimensions of a data set along one edge of a matrix. For example, trade-flow data in the [decompr](http://cran.r-project.org/web/packages/decompr/index.html) and [gvc](http://cran.r-project.org/web/packages/gvc/index.html) package have each country-industry combination occur on each edge of the matrix. 

A fat diagonal matrix looks like this.

```{r}
library(diagonals)
fatdiag(12, steps=3)
```

The workhorse function of this package is the `fatdiag` function, which tries behave similarly to the `diag` function from the `base` package, but then with diagonals being **fat**.

We can also use the function to assign values to the diagonal.

```{r}
( m <- matrix(111, nrow=6, ncol=9) )
fatdiag(m, steps=3) <- 5
m
```

As can be seen from the above example, the blocks and matrices do not have to be square.

The diagonal of a matrix can also be extracted.

```{r}
fatdiag(m, steps=3)
```

We can also specify the size of the block in stead of the number of steps.

```{r}
fatdiag(12, size=4)
```

This also gives us flexibility in terms of having non-square blocks (and consequently matrices).

```{r}
fatdiag(12, size=c(3,4) )
```
