---
title: "Social Network Analysis"
author: "Bastiaan Quast"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A typical example in which `diagonals` can be helpful is Social Network Analysis.
For example, if we use matrices to represent friendship perceptions between individuals, then we need a dyadic matrix.

```{r m1}
# generate a dyadic matrix for 3 individuals
m <- matrix(sample(0:1, 9, replace=TRUE), nrow=3, ncol=3)
m
```

Let says that we want to look at second-order connections (i.e. friends of friends).
If we now want to represent the data from both time period in a single object, we need a 4-dimensional array.
Higher-order arrays are hard to visualise, another way of doing this is by representing two dimensions along each of the two edges of a matrix.
We can do this using the **Knonecker Product** (denoted ⊗), which we can call in `R` using the alias `%x%`.

```{r M}
M <- m %x% m
M
```

Feelings of friendship towards oneself aren't always particularly insightful. We can now use the `diagonals` library to eliminate those.

```{r minus_block_matrix}
# load the library
library(diagonals)

# remove the elements along the diagonal of width 2
minus_block_matrix(M, size=2)
```
